---
bibliography: bibliography.bib
csl: ieee-with-url.csl
---

```{r setup5, include=FALSE}
library(data.table)
library(knitr)
library(multiwayvcov)
library(pwr)
library(stargazer)
opts_chunk$set(echo=FALSE, message=FALSE, warning = FALSE)
stargazer.type <- opts_knit$get('rmarkdown.pandoc.to')
stargazer.type <- ifelse(is.null(stargazer.type), 'text', ifelse('latex' == stargazer.type, 'latex', 'html'))
```

# Results Summary

```{r covariate_relabel5}
df.complete <- read.csv('data/Merged Data.csv')

df.newheader <- c(
  'id', 'mind','energy', 'nature', 'tactics', 'identity', 'experimenter',
  'in.treatment', 'gender', 'age', 'education', 'studying', 'learning',
  'reading', 'raw.score', 'adjusted.score', 'region')

names(df.complete) <- df.newheader

dt.complete <- data.table(df.complete)
```

## Statistical Power

Using the formula for power in the Green and Gerber "Field Experiments" book [@fieldExperiments], the team computed the expected statistical power for the experiment based on the sample.

```{r sample_power}
group.summary <- dt.complete[,
  .(mean = mean(adjusted.score), sd = sd(adjusted.score)),
  by = in.treatment
]

# Calculate power - use formula in text G&G, page 93

group.difference <- abs(group.summary[1, mean] - group.summary[2, mean])
group.sd <- max(group.summary[1, sd], group.summary[2, sd])

pnorm((group.difference * sqrt(75) / (2 * group.sd)) - qnorm(1 - 0.025))
```

Statistical power with 75 subjects was calculated to be 0.133. In other words, the team only has a 13.3% chance of finding a statistically significant treatment effect when such an effect really exists.

Additionally, running the Shapiro-Wilk test on the outcomes indicates that the data is unlikely to be normally distributed, indicating that the power computations may need to be more complicated.

```{r sample_normality}
shapiro.test(df.complete$adjusted.score)
```

## Observations for Control Group

First the team examined the fundamentally unanswerable question:

> Are the Sensing (high-detail) and Intuitive (low-detail) populations different from each other in a reading recall task?

Even though the team cannot answer this question with this experiment, identifying whether the specific populations studied in this experiment differ will provide insight as to whether a heterogeneous treatment effect would be explained as making the scores more different or as making the scores more similar.

The alternate hypothesis predicts that there would be a statistically significant difference between Sensing (high-detail) and Intuition (low-detail) subject scores within the control group.

The corresponding null hypothesis is that there is no statistically significant difference between Sensing (high-detail) and Intuition (low-detail) subject scores within the control group.

```{r control_result}
t.test(
  adjusted.score ~ energy,
  data = subset(df.complete, in.treatment == 'Control'))
```

The t-test indicates that the Intuition (low-detail) subjects in the control group scored higher than Sensing (high-detail) subjects in the control group by 4.3 points, and this has a p-value of 0.105. The team cannot draw any causal interpretation from this result (and it is the opposite of the team's hypothesis), but it can be used to establish a baseline for understanding the heterogeneous treatment effect, should one exist.

The direction of the difference indicates that if the heterogeneous treatment effect is positive for Sensing (high-detail) personality types, the treatment makes the populations more similar, while if the heterogeneous treatment effect is negative for Sensing (high-detail) personality types, the treatment makes the populations more different.

## Results for Treatment Group

The first sub-question for this experiment was as follows:

> Does a study guide accompanying the reading assignment improve reading recall for both personality types?

The alternate hypothesis predicts that there would be a statistically significant difference between control group and the treatment group for both the Sensing (high-detail) subjects and the Intuition (low-detail) subjects.

The corresponding null hypothesis is that there is no statistically significant difference between the treatment and the control group for both the Sensing (high-detail) subjects and the Intuition (low-detail) subjects.

Examining the mean scores suggests that the treatment results in a decrease in the mean score for Intuition (low-detail) subjects and an increase in the mean score for the (high-detail) subjects, though there is high variance in the sample.

```{r treatment_means}
dt.complete[
  order(energy, in.treatment),
  .(mean = mean(adjusted.score), sd = sd(adjusted.score)),
  by = .(energy, in.treatment)
]
```

### Treatment Independent of Personality Types

The first t-test checks the null hypothesis concerning the two groups in the aggregate, which assumes that there is no difference between the control group and the treatment group.

```{r treatment_result_aggregate}
t.test(
  adjusted.score ~ in.treatment,
  data = df.complete)
```

The t-test indicates that the subjects in the treatment group scored _higher_ than the subjects in the control group by 1.7 points, and this has a p-value of 0.330.

### Treatment on Sensing Personality Types

The next t-test checks the null hypothesis concerning Sensing (high-detail) subjects, which assumes there is no difference between the Sensing (high-detail) personality types in the control group and the Sensing (high-detail) personality types in the treatment group.

```{r treatment_result_sensing}
t.test(
  adjusted.score ~ in.treatment,
  data = subset(df.complete, energy == 'Sensing'))
```

The t-test indicates that the Sensing (high-detail) subjects in the treatment group scored _higher_ than Sensing (high-detail) subjects in the control group by 3.8 points, and this has a p-value of 0.120.

### Treatment on Intuitive Personality Types

The next t-test checks the null hypothesis concerning Intuitive (low-detail) subjects, which assumes there is no difference between the Intuitive (low-detail) personality types in the control group and the Intuitive (low-detail) personality types in the treatment group.

```{r treatment_result_intuition}
t.test(
  adjusted.score ~ in.treatment,
  data = subset(df.complete, energy == 'Intuition'))
```

The t-test indicates that the Intuitive (low-detail) subjects in the treatment group scored _lower_ than Intuitive (low-detail) subjects in the control group by 1.6 points, and this has a p-value of 0.472.

## Results for Heterogeneous Treatment Effect

The second sub-question for this experiment was as follows:

> Is the improvement in recall from the study guide different between the two personality types?

Given the opposing directions of the treatment effect on the Sensing (high-detail) personality types and the Intuitive (low-detail) personality types, this gives the research team confidence that there is a heterogeneous treatment effect. The following sections formally test for this effect.

### Baseline Model

This heterogeneous treatment effect can be tested using linear regression. The following baseline model predicts the adjusted score:

$$
\begin{split}
AdjustedScore_i = & \beta_0 +
  \beta_1 Sensing_i +
  \beta_2 InTreatment_i +
  \beta_3 InTreatment_i \times Sensing_i
\end{split}
$$

For a subject $i$, $AdjustedScore_i$ corresponds to the subject's adjusted score from the questionnaire, $Sensing_i$ corresponds to whether the person has a Sensing (high-detail) personality type, and $InTreatment_i$ corresponds to whether the subject was provided with the treatment primer.

For reference purposes, the model will be evaluated against a model leveraging only $InTreatment_i$ and against a model leveraging both $Sensing_i$ and $InTreatment_i$ using `stargazer` [@stargazer]:

```{r causal_model_1, results = 'asis'}
model.1 <- lm(
  adjusted.score ~ in.treatment,
  data = df.complete)

model.2 <- lm(
  adjusted.score ~ energy + in.treatment,
  data = df.complete)

model.3 <- lm(
  adjusted.score ~ energy * in.treatment,
  data = df.complete)

stargazer(
  model.1, model.2, model.3,
  header = FALSE,
  type = stargazer.type,
  title = 'Energy and Treatment',
  dep.var.labels = c('Adjusted Score'),
  covariate.labels = c('Sensing', 'Treatment Primer', 'Sensing x Treatment Primer')
)
```

In this model, the average treatment effect for Intuitive (low-detail) personality types is estimated to be $\hat{\beta}_2 = -1.634$ and the average treatment effect for Sensing (high-detail) personality types is estimated to be $\hat{\beta}_2 + \hat{\beta}_3 = -1.634 + 5.395 = 3.761$.

### Region-Sensitive Model

As noted in the analysis pre-check, four researchers conducted the experiment in three locations around the United States. Consequently there existed the possibility of between group variation, thus making it harder to estimate the average treatment effect precisely.

While the between-group variations by region did not appear to be statistically significant, the team believes it is reasonable to continue accounting for this variation in the model, particularly since the original experimental design attempted to account for this variation.

The following model predicts the adjusted score while also accounting for between-region variation:

$$
\begin{split}
AdjustedScore_i = & \beta_0 +
  \beta_1 Sensing_i +
  \beta_2 InTreatment_i +
  \beta_3 InTreatment_i \times Sensing_i + \\ &
  \beta_4 Delaware_i +
  \beta_5 SouthCarolina_i
\end{split}
$$

For a subject $i$, $Delaware_i$ indicates whether the subject participated in Delaware and $SouthCarolina_i$ indicates whether the subject participated in South Carolina. Having a zero for both indicator variables indicates that the subject participated in California.

This model with clustered standard errors is compared to the baseline model with the $Sensing_i$ and $InTreatment_i$ interaction term using `stargazer` [@stargazer]:

```{r causal_model_2, results = 'asis'}
model.4 <- lm(
  adjusted.score ~ energy * in.treatment + region,
  data = df.complete)

# Use clustered standard errors

model.4.se <- sqrt(diag(cluster.vcov(model.4, ~ region)))

stargazer(
  model.3, model.4,
  se = list(NULL, model.4.se),
  header = FALSE,
  type = stargazer.type,
  title = 'Region, Energy, and Treatment',
  dep.var.labels = c('Adjusted Score'),
  covariate.labels = c(
    'Sensing', 'Treatment Primer',
    'Participant in Delaware',
    'Participant in South Carolina',
    'Sensing x Treatment Primer')
)
```

In this model, the average treatment effect for Intuitive (low-detail) personality types is estimated to be $\hat{\beta}_2 = -1.572$ and the average treatment effect for Sensing (high-detail) personality types is estimated to be $\hat{\beta}_2 + \hat{\beta}_3 = -1.572 + 5.123 = 3.551$.

Additionally, the estimated heterogeneous treatment effect is statistically significant with $\alpha = 0.05$.

### Experimenter-Sensitive Model

The following model predicts the adjusted score while also accounting for between-experimenter variation:

$$
\begin{split}
AdjustedScore_i = & \beta_0 +
  \beta_1 Sensing_i +
  \beta_2 InTreatment_i +
  \beta_3 InTreatment_i \times Sensing_i + \\ &
  \beta_6 Chuck_i +
  \beta_7 Grace_i +
  \beta_8 Minhchau_i
\end{split}
$$

For a subject $i$, $Chuck_i$ indicates whether the subject participated in South Carolina with Chuck as the experimenter, $Grace_i$ indicates whether the subject participated in South Carolina with Grace as the experimenter, and $Minhchau_i$ indicates whether the subject participated in California with Minhchau as the experimenter. Having a zero for all indicator variables indicates that the subject participated in Delaware with Carter as the experimenter.

This model with clustered standard errors is compared to the baseline model with the $Sensing_i$ and $InTreatment_i$ interaction term using `stargazer` [@stargazer]:

```{r causal_model_3, results = 'asis'}
model.5 <- lm(
  adjusted.score ~ energy * in.treatment + experimenter,
  data = df.complete)

# Use clustered standard errors

model.5.se <- sqrt(diag(cluster.vcov(model.5, ~ experimenter)))

stargazer(
  model.3, model.5,
  se = list(NULL, model.5.se),
  header = FALSE,
  type = stargazer.type,
  title = 'Experimenter, Energy, and Treatment',
  dep.var.labels = c('Adjusted Score'),
  covariate.labels = c(
    'Sensing', 'Treatment Primer',
    'Administered by Chuck',
    'Administered by Grace',
    'Administered by Minhchau',
    'Sensing x Treatment Primer')
)
```

In this model, the average treatment effect for Intuitive (low-detail) personality types is estimated to be $\hat{\beta}_2 = -1.591$ and the average treatment effect for Sensing (high-detail) personality types is estimated to be $\hat{\beta}_2 + \hat{\beta}_3 = -1.591 + 5.177 = 3.586$. The estimated heterogeneous treatment effect is statistically significant with $\alpha = 0.05$.

Additionally, the estimated heterogeneous treatment effect is statistically significant with $\alpha = 0.05$.